The Spatial Relationship between Social Media, Digital Connectedness, and Social Isolation in China: Revisting Cyber-Optimism

### 2023-05-29 Social-Isolation-in-China.tex
update

---
title: "Connected or Disconnected?"
author: 
  - Jason Gainous^[Duke Kunshan University, jason.gainous@louisville.edu]
  - Andrew W. MacDonald^[Duke Kunshan University, andrew.macdonald@dukekunshan.edu.cn]
date: '`r format(Sys.time(), "%B %d, %Y")`'
abstract: "The following brief paper is simply the preliminary results of our study. We seek to identify the relationship between social media use and feelings of isolationism in China. This study is based on original date. We conducted a nationwide survey of Chinese citizens conducted in 2015 measuring a range of concepts including those here. The preliminary results are promising. We develop an index of personal isolationism and model this as a function of social media use, connection to the outside digital world, and urbanicity, while controlling for a host of individual level factors. The results suggest the following: 1) Social media use does not have a generalized relationship with isolationism, 2) Those who use VPNs/proxy servers tend to feel less isolated than those who do not, 3) Those living in urban environments tend to feel more isolated, 3) There is a strong positive relationship between social media use and feelings of isolation for those who frequently use VPNs/proxy servers, and 4) The is a strong positive relationship between social media use and feelings of isolation for those living in rural China."
keywords: 
  - social isolation
  - social media
  - China
format: pdf
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(sjPlot)
library(ggpubr)
library(grid)
library(stargazer)
library(scales)
library(papeR)
library(gridExtra)
library(knitr)
library(kableExtra)
library(ggcorrplot)

load("china_isolation_data.RData")

china_isolation_data <- china_original_amanzhol
```

# Social Isolation in China

## Introduction

Research is quite mixed when it comes to identifying the relationship between social media use and feelings of isolationism. Early research suggested that the internet would build social capital, or an interconnectedness between people, and that social media platforms were the ultimate way to bring people together. Since then, research indicates that the effect may be the opposite for many. Social media may encourage social isolation, and ultimately, social media users may feel more disconnected from others than those who do not use social media. We know very little about this phenomenon in China. China presents an interesting case because the internet is bifurcated to some degree -- there is the Chinese internet and the outside internet available via VPNs. Being connected to the outside world may lessen feelings of isolation, or it may increase these feelings because digital exposure to a world is much larger than that in people's immediate environment without the ability to have in-person connection to may make one feel isolated. Interestingly, and somewhat ironically though, it may also matter, too, whether someone lives in a city. Research has suggested that urban living can actually lead to feelings of isolation.

## Introduction

Research is quite mixed when it comes to identifying the relationship between social media use and feelings of isolationism. Early research suggested that the internet would build social capital, or an interconnectedness between people, and that social media platforms were the ultimate way to bring people together. Since then, research indicates that the effect may be the opposite for many. Social media may encourage social isolation, and ultimately, social media users may feel more disconnected from others than those who do not use social media. We know very little about this phenomenon in China. China presents an interesting case because the internet is bifurcated to some degree -- there is the Chinese internet and the outside internet available via VPNs. Being connected to the outside world may lessen feelings of isolation, or it may increase these feelings because digital exposure to a world is much larger than that in people's immediate environment without the ability to have in-person connection to may make one feel isolated. Interestingly, and somewhat ironically though, it may also matter, too, whether someone lives in a city. Research has suggested that urban living can actually lead to feelings of isolation.

## Data

@@

The data collection had two phases. For the first phase, Qualtrics collected and submitted to us a trial run of 286 cases. We were able to check the data for reliability and adjust the instrument before proceeding with the second phase, the final data collection. We only made three adjustments, all of which turned out to be important and central to our analysis. For the trial run, we only asked respondents about pro-government Internet posters and did not ask them about potential net-spy or hostile posters. We amended this oversight and it became central to the analysis that follows. In addition, much of our theory focuses on the impact of Chinese Internet users circumventing government filters on the Internet. We initially had a question about whether citizens had jumped the wall to read sensitive political information (see Appendix A -- Q25), but decided to add whether they had done so for entertainment purposes as well (Q26 - to watch foreign movies, television shows, etc.). Finally, we also included an attention filter question as part of a battery of institutional trust questions where respondents were simply told to select the "None at all" response (Q13H). If they did not select this response, we could assume they were not paying attention to the questions, and the survey ended for these respondents and the data were not collected for them. In the end, the full sample included a total of 2292 respondents. This sample size provides for a roughly ±2 margin of error.

@@

## Summary statistics

### Summary stats table

```{r summarystats}
key.vars <- china_isolation_data %>% 
  select(c(connected, gen_social_med, urbanicity, ses, female, CCP_member, age, jump_wall))

kable(summarize(key.vars, type="numeric")) %>% 
  kable_styling(latex_options="striped")
```

### Correlations

```{r correlations}
cors <- cor(key.vars, use="complete.obs")

ggcorrplot(cors, title="Key variables associated with connectedness", 
           lab=TRUE, type="lower", insig="blank")
```

## Models

```{r models}
mod1 <- lm(connected ~ gen_social_med + urbanicity + ses + female + CCP_member + age + jump_wall, data = china_isolation_data)

mod2 <- lm(connected ~ gen_social_med*urbanicity + ses + female + CCP_member + age + jump_wall, data = china_isolation_data)

mod3 <- lm(connected ~ gen_social_med*jump_wall + urbanicity + ses + female + CCP_member + age, data = china_isolation_data)

tab_model(mod1, mod2, mod3, show.ci = FALSE)
```

## Margin plots

```{r marginplots}
a = plot_model(mod2, type = "pred", title = "", terms = c("gen_social_med", "urbanicity")
               , legend.title = "Urbanicity") +
  scale_color_discrete(labels = c("Village", "Small City", "Mid-sized City", "Suburb", "Big City")) +
  ylab("Feelings of Isolation") +
  xlab("Social Media Use")

b = plot_model(mod3, type = "pred", title = "", terms = c("gen_social_med", "jump_wall")
               , legend.title = "Globally Connected") +
  scale_color_discrete(labels = c("Never", "", "", "A Moderate Amount", "", "", "Very Often")) +
  ylab("Feelings of Isolation") +
  xlab("Social Media Use")

grid.arrange(a, b)
```

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Spatial Relationship between Social Media, Digital Connectedness, and Social Isolation in China: Revisting Cyber-Optimism},
  pdfauthor={Jason Gainous; Andrew W. MacDonald},
  pdfkeywords={social isolation, social media, digital
connectedness, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Spatial Relationship between Social Media, Digital
Connectedness, and Social Isolation in China: Revisting Cyber-Optimism}
\author{Jason Gainous\footnote{University of Louisville,
  jason.gainous@louisville.edu} \and Andrew W. MacDonald\footnote{Duke
  Kunshan University, andrew.macdonald@dukekunshan.edu.cn}}
\date{3/5/23}

\begin{document}
\maketitle
\begin{abstract}
Early research and populuar discourse on the promise of the internet was
optimistic. Many believed it would promote social interconnectedness
and, ultimately, be a positive force in people's lives. Since these
early predictions, research has become much less positive, usually
pointing to the negative aspects of social media. Here we argue that the
digital environment can actually create a space that deters feelings of
social isolationism especially for those living outside cities in
restrictive information environments where citizens are often
disconnected from the outside world. Using data from a nationwide survey
of Chinese citizens we conducted in 2015, we find that the more citizens
digitally connect with others, the more likely they are to move
relationships from online to offliine, and the more they use social
media the less personally isolated they feel. Futher, those living in
urban environments tend to feel more isolated, and this spatial factor
seems to moderate the aforementioned relationships. All three of those
observed relationships are stronger for those in less urban
environments.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\hypertarget{social-isolation-in-china}{%
\section{Social Isolation in China}\label{social-isolation-in-china}}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Research is quite mixed when it comes to identifying the relationship
between social media use and feelings of isolationism. Early research
suggested that the internet would build social capital, or an
interconnectedness between people, and that social media platforms were
the ultimate way to bring people together. Since then, research
indicates that the effect may be the opposite for many. Social media may
encourage social isolation, and ultimately, social media users may feel
more disconnected from others than those who do not use social media. We
know very little about this phenomenon in China. China presents an
interesting case because the internet is bifurcated to some degree --
there is the Chinese internet and the outside internet available via
VPNs. Being isolation to the outside world may lessen feelings of
isolation, or it may increase these feelings because digital exposure to
a world is much larger than that in people's immediate environment
without the ability to have in-person connection to may make one feel
isolated. Interestingly, and somewhat ironically though, it may also
matter, too, whether someone lives in a city. Research has suggested
that urban living can actually lead to feelings of isolation.

\hypertarget{introduction-1}{%
\subsection{Introduction}\label{introduction-1}}

Research is quite mixed when it comes to identifying the relationship
between social media use and feelings of isolationism. Early research
suggested that the internet would build social capital, or an
interconnectedness between people, and that social media platforms were
the ultimate way to bring people together. Since then, research
indicates that the effect may be the opposite for many. Social media may
encourage social isolation, and ultimately, social media users may feel
more disconnected from others than those who do not use social media. We
know very little about this phenomenon in China. China presents an
interesting case because the internet is bifurcated to some degree --
there is the Chinese internet and the outside internet available via
VPNs. Being isolation to the outside world may lessen feelings of
isolation, or it may increase these feelings because digital exposure to
a world is much larger than that in people's immediate environment
without the ability to have in-person connection to may make one feel
isolated. Interestingly, and somewhat ironically though, it may also
matter, too, whether someone lives in a city. Research has suggested
that urban living can actually lead to feelings of isolation.

\hypertarget{data}{%
\subsection{Data}\label{data}}

@@

The data collection had two phases. For the first phase, Qualtrics
collected and submitted to us a trial run of 286 cases. We were able to
check the data for reliability and adjust the instrument before
proceeding with the second phase, the final data collection. We only
made three adjustments, all of which turned out to be important and
central to our analysis. For the trial run, we only asked respondents
about pro-government Internet posters and did not ask them about
potential net-spy or hostile posters. We amended this oversight and it
became central to the analysis that follows. In addition, much of our
theory focuses on the impact of Chinese Internet users circumventing
government filters on the Internet. We initially had a question about
whether citizens had jumped the wall to read sensitive political
information (see Appendix A -- Q25), but decided to add whether they had
done so for entertainment purposes as well (Q26 - to watch foreign
movies, television shows, etc.). Finally, we also included an attention
filter question as part of a battery of institutional trust questions
where respondents were simply told to select the ``None at all''
response (Q13H). If they did not select this response, we could assume
they were not paying attention to the questions, and the survey ended
for these respondents and the data were not collected for them. In the
end, the full sample included a total of 2292 respondents. This sample
size provides for a roughly ±2 margin of error.

@@

\end{document}
